\counterwithout{section}{chapter}

\textbf{January 21$^{\text{st}}$}

Topic summary
\begin{itemize}
    \item Motivation for studying computer architecture
    \item Abstraction layers
    \item Moore's Law -- device and cost curves
    \item Single-core processor performance
    \item Current issues
\end{itemize}

\section{Motivation}
This is a field with job opportunities pog


\section{Processor performance}
Instruction-level parallelism $\Rightarrow$ parallel processing


\section{Current roadblocks}
Dennard scaling -- for a given fixed die area, the power consumption doesn't change (add)
$$addhere$$

Amdahl's law
$$addhere$$

embedded moment


\section{The History of Computer Architecture}

Key themes
\begin{itemize}
    \item Core operations / Turing completeness
    \item Universal Turing Machine (program and data are the same thing)
    \item Abstraction -- decoupling physical implementation and the operations it should be able to do (ISA)
    \item Technology-influenced design choices
\end{itemize}



\subsection{Analog Computers}
$V = IR$
Making a comeback -- fix $R$, and then with input $I$, .

\subsection{Digital Computers}
Everything is digital today

Difference engine -- compute derivatives using a single adder
\begin{align*}
    f(n) &= n^2 + n + 41\\
    d_1(n) &= f(n) - f(n-1) = 2n \\
    d_2(n) &= d_1(n) - d_1(n-1) = 2
\end{align*}

Analytical engine -- the first general-purpose digital computer.
The first general-purpose digital computer.
It was programmable.
It had \textbf{memory}, \textbf{arithmetic}, and \textbf{conditional branching}, meaning that it was Turing Complete.
It was originally in decimal (human-friendly).
It had a 40-digit representation;

Used locking amplification, and it had an anticipating carry.

% TODO: Add definition of Turing Completeness

% Add section on Jacquard Loom - inspiration for Babbage's design

Ada Lovelace

Translated lectures, and she expanded on the capabilities for

The Analytical Engine weaves patterns just as the Jacquard loom weaves (?)


Alan Turing

Turing machine and Turing completeness.
Church-Turing thesis.

Book recs: The Information and I Am a Strange Loop

